---
title: "HW3"
output: html_document
date: "2023-10-03"
---

```{r}
library(tidyverse)
library(mdsr)
library(nycflights13)
library(Lahman)
```

QUESTION 1
Use the Batting, Pitching, and People tables in the Lahman package to answer the following questions:

(a) Name every player in baseball history who has accumu- lated at least 300 home runs (HR) AND at least 300 stolen bases (SB). You can find the first and last name of the player in the People data frame. Join this to your result along with the total home runs and total bases stolen for each of these elite players.
```{r}
#variable for total HR and variable for total SB by playerID
batter_summary <- Batting %>% group_by(playerID) %>% 
  summarize(tHR=sum(HR),tSB=sum(SB)) %>% 
  #Filter total HR and total SB.
  filter(tHR>=300, tSB>300)
#Join with 'people' to show players' names
batter_summary %>% 
  left_join(People %>% select(playerID, nameFirst, nameLast), by = c("playerID" = "playerID"))
```

(b) Similarly, name every pitcher in baseball history who has accumulated at least 300 wins (W) and at least 3,000 strikeouts (SO).
```{r}
#variables to show total W and total SO per playerID
pitcher_summary <- Pitching %>% group_by(playerID) %>% 
  summarize(wins=sum(W), strikeouts=sum(SO)) %>% 
  #Filter total W and total SO.
  filter(wins>=300, strikeouts>=3000)
#Join with 'people' to show players' names.
pitcher_summary %>% 
  left_join(People %>% select(playerID, nameFirst, nameLast), by = c("playerID" = "playerID"))
```

(c) Identify the name and year of every player who has hit at least 50 home runs in a single season. Which player had the lowest batting average in that season?
```{r}
#variables for total HR and batting average per player, per year.
Batting %>% group_by(playerID, yearID) %>% 
  summarize(HR = sum(HR), bat_ave = sum(H)/sum(AB)) %>% 
  #Filter total HR
  filter(HR>=50) %>% 
  #Join with 'people' to get players' names.
  left_join(People %>% select(playerID, nameFirst, nameLast), by = c("playerID" = "playerID")) %>% 
  arrange(bat_ave) %>% head(1)
```
Pete Alonso

QUESTION 2
Use the nycflights13 package and the flights and planes tables to answer the following questions:
```{r}
head(planes)
head(flights)
```

(a) What is the oldest plane (specified by the tailnum vari- able) that flew from New York City airports in 2013?
```{r}
#select relevant columns
old_plane <- flights %>% select(tailnum, origin) %>% 
  #join with selected 'planes' columns
  left_join(planes %>% select(tailnum, year), by = c("tailnum" = "tailnum")) %>% 
  #Group by tailnum and year to ensure each tailnum only appears once in the tailnum column.
  group_by(tailnum,year) %>% 
  summarise(count=n()) %>% 
  #Arrange to get plane with earliest year.
  arrange(year)
head(old_plane,1)
```
N381AA is the oldest plane, created in 1956.

(b) How many airplanes that flew from New York City are included in the planes table?
```{r}
#select relevant columns
incl_plane <- flights %>% select(tailnum, origin) %>% 
  #group by tailnum to ensure each tailnum only appears once in the tailnum column.
  group_by(tailnum) %>% 
  summarise(nyplane = n()) %>% 
  #Join selected 'planes' columns and remove any years with null values.
  inner_join(planes %>% select(tailnum,year), by = c("tailnum"="tailnum")) %>%
  filter(!is.na(year)) %>% 
  arrange("tailnum")

count(incl_plane)
```

QUESTION 3
Convert the following data frame to wide format
```{r}
dat <- data.frame(grp = c("A","A","B","B"),
                  sex = c("F","M","F","M"),
                  meanL = c(0.225,0.47,0.325,0.547),
                  sdL = c(0.106,.325,.106,.308),
                  meanR = c(.34,.57,.4,.647),
                  sdR = c(0.0849, 0.325, 0.0707, 0.274)
                  )
#Use pivot_wider() to convert to a wider format.
dat %>% pivot_wider(names_from = sex, values_from = c(meanL, sdL, meanR,sdR))
```

QUESTION 4
Consider the pccc icd10 dataset.
```{r}
library(pccc)
head(pccc_icd10_dataset)
```

(a) Remove all the columns labeled with ”g” and a number.
```{r}
#select columns with names don't start with g (^g) followed by a number ([0-9])
pccc_new <- select(pccc_icd10_dataset, -matches("^g[0-9]"))
head(pccc_new)
```

(b) Convert the data set from (a) to a long data set with three columns: id, type (pc or dx), and code.
```{r}
#pivot_longer() to convert to longer format.
pccc_new %>% pivot_longer(-id, names_to = "type", values_to = "code")
```

Python Version:
https://colab.research.google.com/drive/1IuNwaOdr2CMAaYrfnZqDl5xEkfEWWqZ2?usp=sharing 
